SULI Report!

### 2015-08-19 doc/suli-report-swe.tex
Added 5 figures and captions to the SULI report. Not all figures are final - a few are mostly for planning the placement of the final figures. #81

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{mathtools}
%\usepackage{multicol}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

%\tableofcontents

@@

\noindent and $r\leq1$ is the ratio of the semi-minor and semi-major axis of the ellipse. This compact notation combines the eccentricity and orientation of the ellipse into a single object. A plot from (Schneider et al.) showing the shape of elliptical galaxies for various values of $\varepsilon_1$ and $\varepsilon_2$ is shown in Figure (??).

%************************
(Galaxy Ellipticity Figure)
%************************

@@

@@

\noindent where an asterisk denotes the complex conjugate\cite{schneider}. The behavior of the distortion relies strongly on the magnitude of $g$; the
effect is called \textit{strong} lensing if $|g|>1$ and \textit{weak} lensing if $|g|<1$. The effects of strong lensing can be quite dramatic, distorting sources into large arcs, multiple images, or even an Einstein ring as shown in Figure (??). While strong lenses are rare as the alignment of the source and foreground mass must be nearly perfect, \textit{all} sources are weakly lensed. The effect is small, usually an ellipticity distortion of only a few percent (cite?), but can be detected locally by averaging the ellipticities of all sources in a small region. As the orientations of the sources should be random, it would be expected that
$$\left<\varepsilon\right>=0.$$

@@

@@

\subsection*{Creating Lightcones}
While ideally all foreground mass in a field would be considered when predicting the weak lensing of a background galaxy, it is computationally prohibitive to do so. Instead, all foreground halos contained within a `lightcone' centered along the line of sight to the source and extending out to a chosen lightcone radius $R$ were considered when calculating the lensing contributions for the background source. A cartoon model of this process is shown in Figure (???). Unless otherwise specified, experiments in this paper used a radius of 2 arcminutes.

%************************
(Pangloss Cartoon Figure)
%************************

@@

The first test of the halo model was with the ellipticity-ellipticity correlation function with a lightcone radius of 2 arcminutes which is given in Figure (??). The statistic measured the average correlation between pairs of ellipticities at separation distances between (0.1??) arcminutes and 2 arcminutes. The correlation function of the intrinsic ellipticities plotted in blue shows no significant deviation from zero as expected, as before lensing the galaxies had random orientations. Both the ray-traced, plotted in green, and halo model, plotted in purple, ellipticities show positive correlation on small scales up to a separation of 0.8 arcminutes. However, the ray-traced correlation is positive on all calculated scales while the halo model does not predict any significant corrlelation at separations above 0.8 arcminutes. Additionally, while the shapes of the correlation functions for both sets of lensed ellipticities aggree well until 0.8 arcminutes, there is a clear systematic underprediction of shear power by the halo model on all scales. On average, the shear power is underpredicted by (??) percent.

%************************
(e-e Correlation Function Plot)
%************************

Using the same catalog of background sources, this statistic was recalculated for the ray-traced and halo model lensed ellipticities for various lightcone radii ranging from $R=1$ arcminute to $R=???$ arcminutes. The result is shown in Figure (??).

%************************
(e-e Correlation Function Plot with multiple R's)
%************************

@@

%************************
(n-g Correlation Function Plot)\\
%************************
(Make a plot of n-g correlation function with multiple R's?? - No, mention that it was made but does not contain any interesting results)

@@

%A simple halo model was used to reconstruct the mass along a line of sight asfd in the Millennium Simulation to make predictions of how the foreground mass weakly lensed the ellipticities of (x \# of bg) of generated backgrond sources. The lensed ellipticities were characterized globally using the ellipticity-ellipticity and galaxy-mass correlation functions and compared to the ray-traced correlations.

>>>>

\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{mathtools}
\usepackage{caption}
\usepackage{subcaption}
%\usepackage{multicol}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

%\tableofcontents
%\onecolumngrid

@@

\noindent and $r\leq1$ is the ratio of the semi-minor and semi-major axis of the ellipse. This compact notation combines the eccentricity and orientation of the ellipse into a single object. A plot from (Schneider et al.) showing the shape of elliptical galaxies for various values of $\varepsilon_1$ and $\varepsilon_2$ is shown in Figure \ref{ellipses}.

@@

%% Ellipticity Orientation Figure
\begin{figure*}
    \centering
    \begin{subfigure}[H]{0.415\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/ellipse_orientations.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The shape of a galaxy image for various ellipticity components $\varepsilon_1$ and $\varepsilon_2$ on the $x$ and $y-$axes respectively. Taken from \cite{schneider}.}
        \label{ellipses}
    \end{subfigure}
    ~
    %% Einstein Ring Figure
    \begin{subfigure}[H]{0.425\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/einstein_ring.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{An image of the luminous red galaxy LRG 3-757 along with a strongy lensed background galaxy, called the `Cosmic Horseshoe'\cite{einstein_ring}}
        \label{einstein_ring}
    \end{subfigure}
    \caption{}
\end{figure*}

@@

\noindent where an asterisk denotes the complex conjugate\cite{schneider}. The behavior of the distortion relies strongly on the magnitude of $g$; the
effect is called \textit{strong} lensing if $|g|>1$ and \textit{weak} lensing if $|g|<1$. The effects of strong lensing can be quite dramatic, distorting sources into large arcs, multiple images, or even an Einstein ring as shown in Figure \ref{einstein_ring}. While strong lenses are rare as the alignment of the source and foreground mass must be nearly perfect, \textit{all} sources are weakly lensed. The effect is small, usually an ellipticity distortion of only a few percent (cite?), but can be detected locally by averaging the ellipticities of all sources in a small region. As the orientations of the sources should be random, it would be expected that
$$\left<\varepsilon\right>=0.$$

@@

%% Pangloss Cartoon
\begin{figure*}
    \centering
    \includegraphics[width=0.85\textwidth]{figs-swe/pangloss_cartoon.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{A cartoon model of how the \texttt{Pangloss} framework uses a dark matter halo model to reconstruct the foreground mass along the line of sight of a background source galaxy. The shear and convergence at the line of sight contributed by each halo is calculated, and the predicted ellipticity is calculated using the sum of these contributions.}
    \label{pangloss_cartoon}
\end{figure*}

@@

\subsection*{Creating Lightcones}

While ideally all foreground mass in a field would be considered when predicting the weak lensing of a background galaxy, it is computationally prohibitive to do so. Instead, all foreground halos contained within a `lightcone' centered along the line of sight to the source and extending out to a chosen lightcone radius $R$ were considered when calculating the lensing contributions for the background source. A cartoon model of this process is shown in Figure \ref{pangloss_cartoon}. Unless otherwise specified, experiments in this paper used a radius of 2 arcminutes.

@@

The first test of the halo model was with the ellipticity-ellipticity correlation function with a lightcone radius of 2 arcminutes which is given in Figure \ref{gg_corr}. The statistic measured the average correlation between pairs of ellipticities at separation distances between (0.1??) arcminutes and 2 arcminutes. The correlation function of the intrinsic ellipticities plotted in blue shows no significant deviation from zero as expected, as before lensing the galaxies had random orientations. Both the ray-traced, plotted in green, and halo model, plotted in purple, ellipticities show positive correlation on small scales up to a separation of 0.8 arcminutes. However, the ray-traced correlation is positive on all calculated scales while the halo model does not predict any significant corrlelation at separations above 0.8 arcminutes. Additionally, while the shapes of the correlation functions for both sets of lensed ellipticities aggree well until 0.8 arcminutes, there is a clear systematic underprediction of shear power by the halo model on all scales. On average, the shear power is underpredicted by (??) percent.

%% Ellipticity-Ellipticity Correlation
\begin{figure}
    \centering
    \begin{subfigure}[h]{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/gg_corr.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The ellipticity-ellipticity correlation function for (x \# of bg) sources at separation distances from 0.1 arcminutes to 2 arcminutes and a lightcone radius of $R=2$ arcminutes. While the correlation of the intrinsic ellitpicities before lensing (blue) shows no correlation, the correlation of the ray-traced ellipticities (green) and halo model predicted ellipticities (purple) both show significant positive correlation on scales up to 0.8 arcminutes. The halo model predicts no correlation above this scale, while the ray-tracing shows positive correlation on all scales.}
        %The correlation of the intrinsic ellitpicities before lensing is plotted in blue, and shows no correlation. The correlation of the true ray-traced ellipticities is plotted in green and shows significant positive correlation on all scales. The halo predicted correlation, plotted in purple, shows significant positive correlation until a separation distance of 0.8 arcminutes.
        \label{gg_corr}
    \end{subfigure}
    ~~
    \begin{subfigure}[h]{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/gg_corr_series.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The ellipticity-ellipticity correlation functions for a series of halo model frameworks with increasing lightcone radii from $R=1$ arminute to $R=8$ arcminutes, all plotted in purple with various line styles. These are compared to the ray-traced correlation using a radius of 2 arcminutes plotted in green.}
        \label{gg_corr_series}
    \end{subfigure}
    \caption{}
\end{figure}

Using the same catalog of background sources, this statistic was recalculated for the ray-traced and halo model lensed ellipticities for various lightcone radii ranging from $R=1$ arcminute to $R=???$ arcminutes. The result is shown in Figure \ref{gg_corr_series}.

@@

%% Galaxy-mass correlation function
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/ng_corr.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{The galaxy-mass correlation function of (x \# of bg) sources at separation distances from 0.1 arcminutes to 2 arcminutes and a lightcone radius of $R=2$ arcminutes. There is no significant correlation of the intrinsic ellipticities (blue) as expected, and both the ray-traced corrleation (green) and halo model correlation (purple) are positive on all scales. The green and purple curves are in agreement from 0.? to 0.? arcminutes, but have significant discrepancies on scales above and below this range.}
    \label{pangloss_cartoon}
\end{figure}
(Make a plot of n-g correlation function with multiple R's?? - No, mention that it was made but does not contain any interesting results)

@@

A simple halo model was used to reconstruct the mass along line of sights in the Millennium Simulation to make predictions of how the foreground mass weakly lensed the ellipticities of (x \# of bg) of generated background sources. The lensed ellipticities were characterized globally using the ellipticity-ellipticity and galaxy-mass correlation functions and compared to the ray-traced correlations.
